---
title: "Emergent Facial Configurations"
author: "Jeffrey Girard"
date: today
format:
  html:
    page-layout: full
    toc: true
    fig-dpi: 300
    df-print: paged
embed-resources: true
---

## Approach

Each coded FACS event is one momentary configuration of one face. Recordings can
contain multiple faces, coded separately by identity, so configurations from
co-visible faces are counted separately and never merged. Exposure is
duration-weighted person-time: two faces visible for one second contribute two
face-seconds.

Correspondence with hypothesized configurations follows the agreed plan:

- **Metric**: Dice coefficient (`facs::compare_to_prototypes`), the match score
  of Le Mau et al. Eq. 1.
- **Unit**: one best-match Dice per configuration (max over all hypothesized
  configurations, agnostic to emotion category). No summarizing within category.
- **Bands**: Haidt & Keltner cutoffs (none <0.2, weak 0.2-0.4, moderate 0.4-0.7,
  strong >=0.7), all reported.
- **Assessability**: hypothesized configurations are reduced to the project
  scheme. Those retaining no in-scheme AUs are *unassessable* and excluded from
  scoring; those losing some AUs are *partially assessable* and flagged, because
  matches against them are scored on a fragment of the hypothesis.
- Reported both unweighted (per distinct configuration) and exposure-weighted.

```{r setup}
#| message: false
library(tidyverse)
library(furrr)
library(gt)
library(facs)
library(datawizard)
```

```{r params}
band_breaks <- c(0, 0.2, 0.4, 0.7, 1.0)
band_labels <- c("none (<0.2)", "weak (0.2-0.4)", "moderate (0.4-0.7)", "strong (>=0.7)")
waves_included <- c(6, 9)
core_cutoff <- 0.01   # >=1% of full-face time, for the core table
```

## Data

```{r data}
info <-
  left_join(
    read_rds("/mnt/datasets/SEEDLingS/2_PROCESSED/session_info.rds"),
    read_rds("facs_times.rds"),
    by = c("subject", "wave")
  ) |>
  filter(!(subject == 18 & wave == 6), wave %in% waves_included) |>
  mutate(
    valid_session  = session_duration - vis - hat,
    valid_fullface = ffa - vis - hat        # includes neutral full-face time
  )

facs <-
  read_rds("/mnt/datasets/SEEDLingS/2_PROCESSED/facs_coding.rds") |>
  left_join(info, by = c("subject", "wave")) |>
  filter(!(subject == 18 & wave == 6), wave %in% waves_included)

project_scheme <- scheme(list(
  occurrence = c(1:2, 4:39),
  intensity  = c(1:2, 4:39),
  asymmetry  = c(1:2, 4:39)
))

protos <-
  prototypes |>
  filter(is.na(emotion_type) | emotion_type != "compound") |>
  mutate(
    emotion = case_when(
      emotion %in% c("anger", "angry")                      ~ "anger",
      emotion %in% c("confused", "confusion")               ~ "confusion",
      emotion %in% c("desire", "desire food", "desire sex") ~ "desire",
      emotion %in% c("disgust", "disgusted")                ~ "disgust",
      emotion %in% c("fear", "fearful")                     ~ "fear",
      emotion %in% c("happy", "happiness", "joy")           ~ "happiness",
      emotion %in% c("sad", "sadness")                      ~ "sadness",
      emotion %in% c("surprise", "surprised")               ~ "surprise",
      .default = emotion
    )
  )

total_session  <- sum(info$valid_session,  na.rm = TRUE)
total_fullface <- sum(info$valid_fullface, na.rm = TRUE)
n_infants      <- n_distinct(info$subject)
```

## Assessability of hypothesized configurations

Reports how many hypothesized configurations the coding scheme could evaluate.
Needed before any claim that an emotion did or did not occur: an emotion whose
configurations are all unassessable cannot appear in the results at all, and its
absence would say nothing about infants' input.

```{r assessability}
proto_status <-
  protos |>
  mutate(
    name     = paste(source, emotion, config_num, sep = "_"),
    full_aus = map_int(code, \(x) length(pull_numcodes(x))),
    kept_aus = map_int(code, \(x) length(intersect(pull_numcodes(x), project_scheme$occurrence))),
    status   = case_when(
      kept_aus == 0        ~ "unassessable",
      kept_aus <  full_aus ~ "partial",
      .default             = "complete"
    )
  ) |>
  select(name, emotion, code, status, full_aus, kept_aus)

# SUPPLEMENT TABLE S1: assessability of the hypothesized set
tab_s1 <-
  proto_status |>
  summarize(.by = status, configs = n(), emotions = n_distinct(emotion)) |>
  mutate(pct = configs / sum(configs)) |>
  arrange(factor(status, levels = c("complete", "partial", "unassessable")))

tab_s1 |>
  gt() |>
  tab_header(title = "Table S1. Assessability of hypothesized configurations") |>
  fmt_percent(columns = pct, decimals = 1) |>
  cols_label(status = "Status", configs = "Configs", emotions = "Emotions",
             pct = "% of configs")

# Emotions that are ENTIRELY unassessable: these cannot occur by construction.
emotions_lost <-
  proto_status |>
  summarize(.by = emotion,
            all_unassessable = all(status == "unassessable"),
            n_configs = n()) |>
  filter(all_unassessable)

cat("Emotions with no assessable configuration:",
    if (nrow(emotions_lost)) paste(emotions_lost$emotion, collapse = ", ") else "none", "\n")
```

## Build configurations

```{r configs}
plan("multisession", workers = 9)

occ_events <-
  facs |>
  mutate(
    occ = future_map_chr(facs_code, \(x) {
      codes <- tidy(coding(x))[[1]]$occurrence
      paste(codes[codes %in% project_scheme$occurrence], collapse = "+")
    }),
    occ = na_if(occ, ""),
    occ = na_if(occ, "NA")
  ) |>
  drop_na(occ, facs_identity, facs_duration) |>
  filter(facs_duration > 0) |>
  mutate(occ = coding(occ))

obs_table <-
  occ_events |>
  summarize(
    .by = occ,
    n_events   = n(),
    n_subjects = n_distinct(subject),                        # infants
    n_faces    = n_distinct(paste(subject, facs_identity)),  # individuals
    duration   = sum(facs_duration)                          # person-time
  ) |>
  mutate(
    n_aus        = str_count(as.character(occ), "\\+") + 1,
    session_pct  = duration / total_session,
    fullface_pct = duration / total_fullface,
    pct_events   = n_events / sum(n_events),
    pct_time     = duration / sum(duration)   # share of expressive exposure
  ) |>
  arrange(desc(duration))

total_exposure <- sum(obs_table$duration)

cat("Events:", nrow(occ_events),
    "| Distinct configurations:", nrow(obs_table),
    "| Infants:", n_infants, "\n")
```

## Dice match scores

```{r dice}
scores <- compare_to_prototypes(
  obs_table$occ,
  scheme  = project_scheme,
  proto   = protos,
  uncoded = "omit",
  output  = "all"
)

obs_table <-
  obs_table |>
  mutate(
    best_dice = map_dbl(scores, \(s) if (length(s)) max(s, na.rm = TRUE) else NA_real_),
    best_match = map2_chr(scores, best_dice, \(s, b) {
      if (is.na(b) || b == 0) return(NA_character_)
      hit <- names(s)[!is.na(s) & s == b]
      ems <- unique(str_remove(str_remove(hit, "^[^_]+_"), "_[^_]+$"))
      paste(sort(ems), collapse = ", ")
    }),
    # Is the best score achieved by any FULLY assessable hypothesized config?
    best_status = map2_chr(scores, best_dice, \(s, b) {
      if (is.na(b) || b == 0) return(NA_character_)
      hit <- names(s)[!is.na(s) & s == b]
      st  <- proto_status$status[match(hit, proto_status$name)]
      if (any(st == "complete", na.rm = TRUE)) "complete" else "partial only"
    }),
    exact = !is.na(best_dice) & best_dice == 1,
    band  = cut(best_dice, breaks = band_breaks, labels = band_labels,
                include.lowest = TRUE, right = FALSE)
  )
```

## FIGURE: match by configuration size and assessability

```{r figure-match}
#| fig-width: 6.5
#| fig-height: 7.0

fig_match <-
  obs_table |>
  filter(!is.na(best_dice), !is.na(best_status)) |>
  mutate(
    complexity = factor(
      pmin(n_aus, 5),
      levels = 1:5,
      labels = c("1 AU", "2 AUs", "3 AUs", "4 AUs", "5+ AUs")
    ),
    best_status = factor(
      best_status,
      levels = c("complete", "partial only"),
      labels = c("Fully assessable", "Partially assessable")
    )
  ) |>
  summarize(.by = c(best_dice, complexity, best_status), w = sum(duration)) |>
  mutate(p = w / total_exposure) |>
  ggplot(aes(x = best_dice, y = p)) +
  geom_vline(xintercept = 0.7, linetype = "dashed", color = "darkorange2") +
  geom_segment(aes(xend = best_dice, yend = 0), linewidth = 0.6) +
  geom_point(aes(shape = best_status), size = 1.8, fill = "white",
             show.legend = FALSE, stroke = 1) +
  scale_shape_manual(values = c(16, 21)) +
  facet_grid(complexity ~ best_status) +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.25)) +
  scale_y_continuous(labels = scales::percent) +
  coord_cartesian(ylim = c(0, 0.175)) +
  labs(x = "Best-match Dice coefficient",
       y = "Share of infants' facial exposure (person-time)") +
  theme_bw() +
  theme(panel.grid.minor = element_blank())

fig_match
ggsave("fig_match.png", fig_match, width = 6.5, height = 7.0, dpi = 300)
```

## MAIN TABLE: core configurations

```{r table-core}
tab_core <-
  obs_table |>
  filter(fullface_pct >= core_cutoff) |>
  transmute(
    occ = as.character(occ), n_aus, n_events, n_subjects, n_faces,
    duration, fullface_pct, session_pct, best_dice,
    best_match = coalesce(best_match, "\u2014")
  )

tail_row <-
  obs_table |>
  filter(fullface_pct < core_cutoff) |>
  summarize(
    occ = sprintf("Remaining %d configurations (each <1%%)", n()),
    n_aus = NA_real_, n_events = sum(n_events),
    n_subjects = NA_integer_, n_faces = NA_integer_,
    duration = sum(duration),
    fullface_pct = sum(duration) / total_fullface,
    session_pct = sum(duration) / total_session,
    best_dice = NA_real_, best_match = ""
  )

bind_rows(tab_core, tail_row) |>
  gt() |>
  tab_header(
    title = "Emergent facial configurations in infants' visual input",
    subtitle = sprintf("Configurations occupying >=1%% of full-face time (%d of %d)",
                       nrow(tab_core), nrow(obs_table))
  ) |>
  fmt_number(columns = c(duration, best_dice), decimals = 2) |>
  fmt_percent(columns = c(fullface_pct, session_pct), decimals = 2) |>
  fmt_integer(columns = c(n_aus, n_events, n_subjects, n_faces)) |>
  sub_missing(missing_text = "\u2014") |>
  cols_label(
    occ = "AUs", n_aus = "#AU", n_events = "Events",
    n_subjects = paste0("# Infants (of ", n_infants, ")"),
    n_faces = "# Individuals", duration = "Duration (s)",
    fullface_pct = "% Full-face time", session_pct = "% Session time",
    best_dice = "Best Dice", best_match = "Best-matching emotion(s)"
  ) |>
  tab_source_note(md(
    "Duration is person-time; co-visible faces contribute independently. *# Individuals* is distinct infant-by-identity combinations. *Best-matching emotion(s)* lists every emotion achieving the best Dice; a configuration matching several reflects overlap among the hypothesized sets and these are not additive."
  ))
```

## MAIN TABLE: match distribution

```{r table-bands}
tab_bands <-
  obs_table |>
  filter(!is.na(band)) |>
  summarize(.by = band, configs = n(), events = sum(n_events), seconds = sum(duration)) |>
  mutate(
    band         = factor(as.character(band),
                          levels = c(band_labels, "no configuration coded")),
    pct_configs  = configs / sum(configs),
    pct_fullface = seconds / total_fullface,
    pct_session  = seconds / total_session
  )

neutral_seconds <- total_fullface - sum(obs_table$duration)

neutral_row <- tibble(
  band         = factor("no configuration coded",
                        levels = c(band_labels, "no configuration coded")),
  configs      = NA_integer_,
  events       = NA_integer_,
  seconds      = neutral_seconds,
  pct_configs  = NA_real_,
  pct_fullface = neutral_seconds / total_fullface,
  pct_session  = neutral_seconds / total_session
)

bind_rows(tab_bands, neutral_row) |>
  arrange(band) |>
  gt() |>
  tab_header(
    title = "Match to hypothesized configurations",
    subtitle = "Best-match Dice per configuration, by Haidt & Keltner band"
  ) |>
  fmt_number(columns = seconds, decimals = 1) |>
  fmt_percent(columns = c(pct_configs, pct_fullface, pct_session), decimals = 1) |>
  sub_missing(missing_text = "\u2014") |>
  cols_label(
    band = "Band",
    configs = "Configs",
    events = "Events",
    seconds = "Seconds",
    pct_configs = "% of configs (unweighted)",
    pct_fullface = "% of full-face time",
    pct_session = "% of session time"
  ) |>
  tab_source_note(md(
    "Session percentages do not sum to 100%: the remainder is time with no fully visible face."
  ))
```

## IN-TEXT NUMBERS

Everything the results paragraphs need, printed once so nothing is transcribed
by hand.

```{r in-text}
list(
  n_events            = nrow(occ_events),
  n_configs           = nrow(obs_table),
  n_infants           = n_infants,
  # concentration
  n_core              = sum(obs_table$fullface_pct >= core_cutoff),
  core_pct_time       = sum(obs_table$duration[obs_table$fullface_pct >= core_cutoff]) / total_exposure,
  # expressiveness: how much available face time was expressive at all
  expressive_pct_fullface = total_exposure / total_fullface,
  expressive_pct_session  = total_exposure / total_session,
  # spread
  pct_configs_one_infant  = mean(obs_table$n_subjects == 1),
  median_infants          = median(obs_table$n_subjects),
  # exact matches
  exact_configs       = sum(obs_table$exact, na.rm = TRUE),
  exact_pct_configs   = mean(obs_table$exact, na.rm = TRUE),
  exact_pct_time      = sum(obs_table$duration[obs_table$exact], na.rm = TRUE) / total_exposure,
  # assessability of the matches
  partial_only_pct_time = sum(obs_table$duration[obs_table$best_status == "partial only"], na.rm = TRUE) / total_exposure
) |>
  enframe(name = "quantity", value = "value") |>
  mutate(value = map_chr(value, \(x) format(unlist(x), digits = 3))) |>
  print(n = Inf)

describe_distribution(obs_table, select = "best_dice")
```

```{r in-text-complexity}
# The complexity gradient: the figure's punchline as a number.
obs_table |>
  filter(!is.na(best_dice)) |>
  summarize(
    .by = n_aus,
    configs = n(),
    seconds = sum(duration),
    exact_seconds = sum(duration[exact])
  ) |>
  mutate(
    pct_of_exposure = seconds / total_exposure,
    pct_of_exact    = exact_seconds / sum(exact_seconds)
  ) |>
  arrange(n_aus) |>
  gt() |>
  tab_header(title = "Exposure and exact matches by configuration size") |>
  fmt_number(columns = c(seconds, exact_seconds), decimals = 1) |>
  fmt_percent(columns = c(pct_of_exposure, pct_of_exact), decimals = 1) |>
  cols_label(n_aus = "AUs in config", configs = "Configs", seconds = "Seconds",
             exact_seconds = "Exact-match sec", pct_of_exposure = "% exposure",
             pct_of_exact = "% of exact-match time")
```

```{r}
#| output: asis
tk <- list(
  n_configs   = nrow(obs_table),
  n_core      = sum(obs_table$fullface_pct >= core_cutoff),
  core_time   = sum(obs_table$duration[obs_table$fullface_pct >= core_cutoff]) / total_exposure,
  expressive  = total_exposure / total_fullface,
  one_infant  = mean(obs_table$n_subjects == 1),
  exact_time  = sum(obs_table$duration[obs_table$exact], na.rm = TRUE) / total_exposure,
  small_exact = sum(obs_table$duration[obs_table$exact & obs_table$n_aus <= 2], na.rm = TRUE) /
                sum(obs_table$duration[obs_table$exact], na.rm = TRUE),
  big_time    = sum(obs_table$duration[obs_table$n_aus >= 5]) / total_exposure
)

cat(sprintf(
"## Key takeaways

1. **Faces are usually still.** Only %.1f%% of full-face time had any coded AU activity.
2. **A few configurations dominate.** %d configurations (of %d) each occupy >=1%% of full-face time and together account for %.1f%% of expressive exposure.
3. **The tail is idiosyncratic.** %.0f%% of configurations occurred in only one infant's input.
4. **Matches are common but simple.** %.1f%% of expressive exposure exactly matches a hypothesized configuration, but %.0f%% of that exact-match time is configurations of one or two AUs.
5. **Elaborate configurations are near-absent.** Configurations of 5+ AUs account for %.1f%% of expressive exposure, so the multi-AU configs typically depicted in the emotion literature are very rare in infants' input.
",
  100 * tk$expressive, tk$n_core, tk$n_configs, 100 * tk$core_time,
  100 * tk$one_infant, 100 * tk$exact_time, 100 * tk$small_exact, 100 * tk$big_time
))
```


## SUPPLEMENT

```{r supplement}
# S2: full catalogue (all configurations)
obs_table |>
  transmute(occ = as.character(occ), n_aus, n_events, n_subjects, n_faces,
            duration, fullface_pct, session_pct, best_dice, best_status,
            best_match, band) |>
  write_csv("supp_S2_full_catalogue.csv")

# S3: every exact match, with the size of the config and its matched emotions
obs_table |>
  filter(exact) |>
  arrange(desc(duration)) |>
  transmute(occ = as.character(occ), n_aus, n_events, duration,
            fullface_pct, n_subjects, best_status, best_match) |>
  write_csv("supp_S3_exact_matches.csv")

# S4: assessability detail, per hypothesized configuration
proto_status |>
  transmute(name, emotion, code = as.character(code), full_aus, kept_aus, status) |>
  write_csv("supp_S4_assessability.csv")

write_rds(obs_table, "obs_table.rds")
```

## What goes where

**Main text figure**: `fig_match.png`. Best-match Dice by configuration size and
assessability. Caption must state that heights are shares of *total* facial
exposure across all ten panels (so columns do not each sum to 100%), and define
assessability.

**Main text tables**: core configurations, and the band distribution.

**In text, not tabled**: total configurations; the fraction of available
full-face time that was expressive at all (this is the "faces are mostly doing
nothing" point); the spread numbers; the exact-match rate; and the complexity
gradient (single- and two-AU configurations account for most exact-match time,
while 5+ AU configurations are near-absent).

**Supplement**: S1 assessability summary, S2 full catalogue, S3 exact matches,
S4 per-configuration assessability.

**Deliberately not reported**: per-emotion occurrence counts. Configurations
match multiple emotions because the hypothesized sets overlap, so per-emotion
counts are not additive and invite misreading. Emotion labels appear only as a
descriptive column in the core table.
