---
title: "Face Presence"
author: "Jeffrey Girard"
date: today
format:
  html:
    page-layout: full
    toc: true
    fig-width: 6.5
    fig-height: 4
    fig-dpi: 300
    df-print: paged
embed-resources: true
---

## Approach

How much of an infant's recorded day contained a visible face?

- **Unit**: the video timeline. Face presence is a property of the infant's view,
  not of any individual person, so two co-visible faces do not make the view
  twice as face-filled. Durations are therefore a **wall-clock union**: intervals
  are merged before summing.
- This differs deliberately from the FACS configuration analyses, which use
  person-time, because a *configuration* belongs to a face while *presence* does
  not.
- **Denominator**: total session duration (not `valid_fullface`). The question is
  what fraction of the recorded day had a face, so occluded and face-free time
  both belong in the base.

```{r setup}
#| message: false
library(tidyverse)
library(brms)
library(easystats)
library(ggeffects)
library(gt)
```

```{r data}
info <- read_rds("/mnt/datasets/SEEDLingS/2_PROCESSED/session_info.rds")

part <- read_rds("/mnt/datasets/SEEDLingS/2_PROCESSED/partface_coding.rds") |>
  left_join(info, by = join_by(subject, wave, month))

full <- read_rds("/mnt/datasets/SEEDLingS/2_PROCESSED/fullface_coding.rds") |>
  left_join(info, by = join_by(subject, wave, month))
```

## Union durations

```{r union-fn}
# Merge overlapping intervals within a session, then sum. Face presence is a
# property of the timeline, so co-visible faces must not double-count.
union_duration <- function(df, onset, offset) {
  df |>
    arrange(subject, wave, month, {{ onset }}) |>
    mutate(
      .by = c(subject, wave, month),
      .max_off = cummax(lag({{ offset }}, default = -Inf)),
      .new_blk = {{ onset }} > .max_off,
      .blk     = cumsum(.new_blk)
    ) |>
    summarize(
      .by = c(subject, wave, month, .blk),
      .on  = min({{ onset }}),
      .off = max({{ offset }}),
      session_duration = first(session_duration)
    ) |>
    summarize(
      .by = c(subject, wave, month),
      sum_sec = sum(.off - .on),
      session_duration = first(session_duration)
    )
}
```

```{r build}
# window is 0-based in BOTH datasets. The previous version used 0-based for
# partial and 1-based for full, so the two models' intercepts were not
# comparable and the full-face conditional-effects plot requested a window value
# that did not exist in its data.
df_part <-
  part |>
  filter(part_duration > 0) |>
  union_duration(part_onset, part_offset) |>
  mutate(
    pct_part = sum_sec / session_duration,
    window   = match(wave, c(6, 9, 12, 15)) - 1   # wave -> 0,1,2,3
  ) |>
  complete(subject, window)

df_full <-
  full |>
  filter(full_duration > 0) |>
  union_duration(full_onset, full_offset) |>
  mutate(pct_full = sum_sec / session_duration) |>
  arrange(subject, wave) |>
  mutate(.by = subject, window = row_number() - 1) |>
  complete(subject, window)
```

## MAIN TABLE: face presence by age window

```{r table-descriptives}
bind_rows(
  df_part |> drop_na(pct_part) |>
    summarize(.by = wave, n = n(),
              Mean = mean(pct_part), SD = sd(pct_part),
              Median = median(pct_part), MAD = mad(pct_part)) |>
    mutate(Measure = "Partial face"),
  df_full |> drop_na(pct_full) |>
    summarize(.by = wave, n = n(),
              Mean = mean(pct_full), SD = sd(pct_full),
              Median = median(pct_full), MAD = mad(pct_full)) |>
    mutate(Measure = "Full face")
) |>
  mutate(
    Window = factor(wave, levels = c(6, 9, 12, 15),
                    labels = c("6-8 mo", "9-11 mo", "12-14 mo", "15-17 mo"))
  ) |>
  select(Measure, Window, n, Mean, SD, Median, MAD) |>
  arrange(desc(Measure), Window) |>
  gt(groupname_col = "Measure") |>
  tab_header(
    title = "Face presence in infants' recorded day",
    subtitle = "Share of session duration with a face visible"
  ) |>
  fmt_percent(columns = c(Mean, SD, Median, MAD), decimals = 1) |>
  tab_options(row_group.as_column = TRUE)
```

## Models

Beta regression on the proportion of session time with a face, with a random
intercept per infant. `mi()` handles the sessions missing from the complete grid.

```{r models}
#| warning: false

fit_presence <- function(data, outcome, form, file) {
  brm(
    formula = form,
    family  = Beta(link = "logit", link_phi = "log"),
    data    = data,
    prior   = if (str_detect(as.character(form)[3], "window"))
                set_prior("normal(0, 0.5)", class = "b") else NULL,
    iter = 10000, warmup = 9000, chains = 4, cores = 4,
    refresh = 0, init = 0, file = file,
    control = list(adapt_delta = 0.99, max_treedepth = 12),
    backend = "cmdstanr",
  )
}

part_null      <- fit_presence(df_part, form = bf(pct_part | mi() ~ 1 + (1 | subject)),
                               file = "part_null_union")
part_linear    <- fit_presence(df_part, form = bf(pct_part | mi() ~ 1 + window + (1 + window | subject)),
                               file = "part_linear_union")
part_monotonic <- fit_presence(df_part, form = bf(pct_part | mi() ~ 1 + mo(window) + (1 + mo(window) | subject)),
                               file = "part_monotonic_union")

full_null      <- fit_presence(df_full, form = bf(pct_full | mi() ~ 1 + (1 | subject)),
                               file = "full_null_union")
full_linear    <- fit_presence(df_full, form = bf(pct_full | mi() ~ 1 + window + (1 + window | subject)),
                               file = "full_linear_union")
full_monotonic <- fit_presence(df_full, form = bf(pct_full | mi() ~ 1 + mo(window) + (1 + mo(window) | subject)),
                               file = "full_monotonic_union")
```

## MAIN TABLE: change over age

```{r table-results}
res <-
  list("Partial face" = part_linear, "Full face" = full_linear) |>
  map(\(m) {
    describe_posterior(m, parameters = "window") |>
      as_tibble() |>
      transmute(
        Est. = Median, CL = CI_low, CU = CI_high,
        OR = exp(Est.), OR_CL = exp(CL), OR_CU = exp(CU),
        pval = 2 * (1 - pd)
      )
  }) |>
  list_rbind(names_to = "Outcome")

gt(res) |>
  tab_header(title = "Change in face presence across age windows") |>
  fmt_number(columns = Est.:OR_CU, decimals = 2) |>
  fmt(columns = pval, fns = \(x) {
    cx <- sub("^0\\.", ".", sprintf("%.3f", x))
    ifelse(x < 0.001, "<.001", cx)
  }) |>
  cols_merge(columns = c(Est., CL, CU), pattern = "{1} ({2}, {3})") |>
  cols_merge(columns = c(OR, OR_CL, OR_CU), pattern = "{1} ({2}, {3})") |>
  cols_label(Est. = "Logit scale", OR = "Odds ratio", pval = "p-value") |>
  cols_align(align = "center", columns = c(Est., OR))
```

## FIGURE: face presence across age

```{r figure}
#| fig-width: 6.5
#| fig-height: 4
#| message: false

ce_df <- function(model, label) {
  plot(conditional_effects(model, effects = "window",
                           int_conditions = list(window = 0:3)), plot = FALSE)[[1]]$data |>
    as_tibble() |>
    transmute(window, estimate__, lower__, upper__, Measure = label)
}

fig_presence <-
  bind_rows(ce_df(part_linear, "Partial face"),
            ce_df(full_linear, "Full face")) |>
  ggplot(aes(x = window, y = estimate__, ymin = lower__, ymax = upper__)) +
  geom_ribbon(alpha = 0.15) +
  geom_line() +
  geom_point() +
  facet_wrap(~Measure, scales = "free_y") +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(
    breaks = 0:3,
    labels = c("6-8", "9-11", "12-14", "15-17")
  ) +
  labs(x = "Age window (months)",
       y = "Share of session with a face visible") +
  theme_bw(base_size = 10) +
  theme(panel.grid.minor = element_blank())

fig_presence
ggsave("fig_presence.png", fig_presence, width = 6.5, height = 4, dpi = 300)
```

## IN-TEXT NUMBERS

```{r in-text}
vd_part <- variance_decomposition(part_null)
vd_full <- variance_decomposition(full_null)

list(
  n_sessions_part   = sum(!is.na(df_part$pct_part)),
  n_sessions_full   = sum(!is.na(df_full$pct_full)),
  n_infants         = n_distinct(df_part$subject),
  part_mean         = mean(df_part$pct_part, na.rm = TRUE),
  part_median       = median(df_part$pct_part, na.rm = TRUE),
  full_mean         = mean(df_full$pct_full, na.rm = TRUE),
  full_median       = median(df_full$pct_full, na.rm = TRUE),
  part_icc          = vd_part$ICC_decomposed,
  full_icc          = vd_full$ICC_decomposed
) |>
  enframe(name = "quantity", value = "value") |>
  mutate(value = map_chr(value, \(x) format(unlist(x), digits = 3))) |>
  print(n = Inf)
```

```{r takeaways}
#| output: asis

cat(sprintf(
"## Key takeaways

1. **Faces are rare in infants' view.** A full face was visible for a median of
   %.1f%% of the session; a partial face for %.1f%%.
2. **Infants differ from one another.** The intercept-only models attribute
   %.0f%% (partial) and %.0f%% (full) of the variance to stable between-infant
   differences.
",
  100 * median(df_full$pct_full, na.rm = TRUE),
  100 * median(df_part$pct_part, na.rm = TRUE),
  100 * vd_part$ICC_decomposed,
  100 * vd_full$ICC_decomposed
))
```

## SUPPLEMENT

### S1. Distributions

```{r supp-hist}
#| fig-width: 6.5
#| fig-height: 3

bind_rows(
  df_part |> transmute(pct = pct_part, Measure = "Partial face"),
  df_full |> transmute(pct = pct_full, Measure = "Full face")
) |>
  drop_na(pct) |>
  ggplot(aes(x = pct)) +
  geom_histogram(color = "white", breaks = seq(0, 1, 0.025)) +
  facet_wrap(~Measure, scales = "free_y") +
  scale_x_continuous(breaks = seq(0, 1, 0.2), labels = scales::percent) +
  labs(x = "Share of session with a face visible", y = "Sessions") +
  theme_bw(base_size = 10)
```

### S2. Monotonic models

The linear models assume evenly spaced age effects. The monotonic models relax
that, allowing the spacing between windows to be estimated.

```{r supp-monotonic}
model_parameters(part_monotonic, effects = "fixed") |> print_md()
model_parameters(full_monotonic, effects = "fixed") |> print_md()
```

### S3. Model checks

```{r supp-ppc}
#| warning: false
#| fig-width: 6.5
#| fig-height: 3

pp_check(part_linear, ndraws = 50)
pp_check(full_linear, ndraws = 50)
```

### S4. Variance decomposition

```{r supp-vd}
vd_part
vd_full
```
