---
title: "Hypothesized Facial Configurations"
author: "Jeffrey Girard"
date: today
format:
  html:
    page-layout: full
    toc: true
    fig-dpi: 300
    df-print: paged
embed-resources: true
---

## Approach

- **Unit**: each coded FACS event is one momentary configuration of one face.
  Recordings contain multiple faces, coded separately by identity.
- **Exposure**: duration-weighted **person-time**. Overlapping events are merged
  *within* a face (a face cannot double-count itself), then summed *across*
  faces (two co-visible faces contribute independently). The previous version
  merged across faces, taking the wall-clock union, which undercounted exposure
  whenever two people were visible at once.
- **Denominators**: valid full-face time (which includes time a face was visible
  but no AU was coded) and valid session time. Never the share of expressive
  time, which would make faces appear more expressive than they are.
- **Assessability**: hypothesized configurations are reduced to the project
  scheme. Those retaining no in-scheme AUs are *unassessable* and cannot occur
  by construction; those losing some AUs are *partially assessable*, and matches
  against them are scored on a fragment of the hypothesis.
- **Phases are never pooled.** Phase I (waves 6, 9) is the main analysis and is
  the window the emergent analysis also covers. Phase II (waves 12, 15) is
  reported in the supplement, where the caregiver-worn camera makes the two
  windows less comparable.

```{r setup}
#| message: false
library(tidyverse)
library(furrr)
library(gt)
library(facs)
library(easystats)
```

## Data

```{r data}
info <-
  left_join(
    read_rds("/mnt/datasets/SEEDLingS/2_PROCESSED/session_info.rds"),
    read_rds("facs_times.rds"),
    by = c("subject", "wave")
  ) |>
  filter(!(subject == 18 & wave == 6)) |>
  mutate(
    valid_session  = session_duration - vis - hat,
    valid_fullface = ffa - vis - hat,
    phase = if_else(wave %in% c(6, 9), "I", "II")
  )

facs <-
  read_rds("/mnt/datasets/SEEDLingS/2_PROCESSED/facs_coding.rds") |>
  left_join(info, by = c("subject", "wave")) |>
  filter(!(subject == 18 & wave == 6)) |>
  drop_na(facs_identity, facs_duration) |>
  filter(facs_duration > 0) |>
  # identity belongs in the event key: facs_id may restart per face
  mutate(event_id = str_c(subject, wave, facs_identity, facs_id, sep = "_"))

project_scheme <- scheme(list(
  occurrence = c(1:2, 4:39),
  intensity  = c(1:2, 4:39),
  asymmetry  = c(1:2, 4:39)
))

protos <-
  prototypes |>
  filter(is.na(emotion_type) | emotion_type != "compound") |>
  mutate(
    emotion = case_when(
      emotion %in% c("anger", "angry")                      ~ "anger",
      emotion %in% c("confused", "confusion")               ~ "confusion",
      emotion %in% c("desire", "desire food", "desire sex") ~ "desire",
      emotion %in% c("disgust", "disgusted")                ~ "disgust",
      emotion %in% c("fear", "fearful")                     ~ "fear",
      emotion %in% c("happy", "happiness", "joy")           ~ "happiness",
      emotion %in% c("sad", "sadness")                      ~ "sadness",
      emotion %in% c("surprise", "surprised")               ~ "surprise",
      .default = emotion
    )
  )

phase_totals <-
  info |>
  summarize(
    .by = phase,
    valid_session  = sum(valid_session,  na.rm = TRUE),
    valid_fullface = sum(valid_fullface, na.rm = TRUE),
    n_infants      = n_distinct(subject)
  ) |>
  arrange(phase)

phase_totals
```

## Assessability of the hypothesized set

An emotion whose configurations all fall outside the coding scheme cannot occur
in these data. Reporting it as absent would describe our scheme, not infants'
input, so these are excluded rather than reported as zero.

```{r assessability}
proto_status <-
  protos |>
  mutate(
    name     = paste(source, emotion, config_num, sep = "_"),
    full_aus = map_int(code, \(x) length(pull_numcodes(x))),
    kept_aus = map_int(code, \(x) length(intersect(pull_numcodes(x), project_scheme$occurrence))),
    status   = case_when(
      kept_aus == 0        ~ "unassessable",
      kept_aus <  full_aus ~ "partial",
      .default             = "complete"
    ),
    reduced  = map_chr(code, \(x) {
      keep <- sort(intersect(pull_numcodes(x), project_scheme$occurrence))
      if (length(keep)) paste(keep, collapse = "+") else NA_character_
    })
  )

proto_status |>
  summarize(.by = status, configs = n(), emotions = n_distinct(emotion)) |>
  mutate(pct = configs / sum(configs)) |>
  arrange(factor(status, levels = c("complete", "partial", "unassessable"))) |>
  gt() |>
  tab_header(title = "Assessability of hypothesized configurations") |>
  fmt_percent(columns = pct, decimals = 1) |>
  cols_label(status = "Status", configs = "Configs", emotions = "Emotions",
             pct = "% of configs")

emotions_lost <-
  proto_status |>
  summarize(.by = emotion, all_unassessable = all(status == "unassessable")) |>
  filter(all_unassessable) |>
  pull(emotion)

cat("Emotions with no assessable configuration:",
    if (length(emotions_lost)) paste(emotions_lost, collapse = ", ") else "none", "\n")

protos_assessable <-
  proto_status |>
  filter(status != "unassessable") |>
  distinct(emotion, reduced, status) |>
  mutate(occ = coding(reduced))
```

## Observed configurations

```{r occ}
plan("multisession", workers = 9)

occ_events <-
  facs |>
  mutate(
    occ = future_map_chr(facs_code, \(x) {
      codes <- tidy(coding(x))[[1]]$occurrence
      paste(codes[codes %in% project_scheme$occurrence], collapse = "+")
    }),
    occ = na_if(occ, ""),
    occ = na_if(occ, "NA")
  ) |>
  drop_na(occ) |>
  mutate(occ = coding(occ))

occ_events |> count(phase, name = "events")
```

## Exposure (person-time)

```{r exposure}
# Merge overlapping events WITHIN a face, then sum ACROSS faces. Including
# facs_identity in the keys is what prevents a wall-clock union across people.
merge_within_face <- function(df, keys) {
  df |>
    arrange(across(all_of(keys)), facs_onset) |>
    mutate(
      .by = all_of(keys),
      max_off  = cummax(lag(facs_offset, default = -Inf)),
      new_blk  = facs_onset > max_off,
      block_id = cumsum(new_blk)
    ) |>
    summarize(
      .by = c(all_of(keys), block_id),
      onset  = min(facs_onset),
      offset = max(facs_offset)
    ) |>
    mutate(block_dur = offset - onset) |>
    summarize(.by = all_of(keys), duration = sum(block_dur))
}

emotion_events <-
  occ_events |>
  inner_join(
    protos_assessable |> select(emotion, occ),
    by = join_by(occ),
    relationship = "many-to-many"
  ) |>
  # an event may match several variants of one emotion: count it once per emotion
  distinct(subject, wave, phase, facs_identity, emotion, event_id,
           facs_onset, facs_offset, facs_duration)

# subject x wave x emotion person-time (merged within face, summed across faces)
emotion_subject <-
  emotion_events |>
  merge_within_face(keys = c("subject", "wave", "phase", "facs_identity", "emotion")) |>
  summarize(.by = c(subject, wave, phase, emotion), duration = sum(duration))

# complete the grid so non-occurrences are explicit zeros
emotion_long <-
  crossing(
    info |> select(subject, wave, phase, valid_session, valid_fullface),
    emotion = sort(unique(protos_assessable$emotion))
  ) |>
  left_join(emotion_subject, by = join_by(subject, wave, phase, emotion)) |>
  replace_na(list(duration = 0)) |>
  mutate(
    fullface_pct = duration / valid_fullface,
    session_pct  = duration / valid_session
  )

# phase-level summary, with spread across infants
emotion_phase <-
  emotion_long |>
  summarize(
    .by = c(phase, emotion),
    seconds        = sum(duration),
    infants_any    = sum(duration > 0),                 # infants with any exposure
    max_infant_sec = max(duration),                     # largest single contributor
  ) |>
  left_join(phase_totals, by = "phase") |>
  mutate(
    fullface_pct = seconds / valid_fullface,
    session_pct  = seconds / valid_session,
    # concentration: what share of this emotion's exposure comes from one infant
    top_infant_share = if_else(seconds > 0, max_infant_sec / seconds, NA_real_)
  )
```

## MAIN TABLE (Phase I): exposure to hypothesized configurations

```{r table-phase1}
n_inf_i <- phase_totals$n_infants[phase_totals$phase == "I"]

emotion_phase |>
  filter(phase == "I") |>
  arrange(desc(fullface_pct)) |>
  select(emotion, seconds, fullface_pct, session_pct, infants_any, top_infant_share) |>
  mutate(emotion = str_to_title(emotion)) |>
  gt() |>
  tab_header(
    title = "Phase I: exposure to hypothesized facial configurations",
    subtitle = "Person-time; co-visible faces contribute independently"
  ) |>
  fmt_number(columns = seconds, decimals = 1) |>
  fmt_percent(columns = c(fullface_pct, session_pct, top_infant_share), decimals = 2) |>
  sub_missing(missing_text = "\u2014") |>
  cols_label(
    emotion = "Emotion",
    seconds = "Seconds",
    fullface_pct = "% Full-face time",
    session_pct = "% Session time",
    infants_any = paste0("# Infants (of ", n_inf_i, ")"),
    top_infant_share = "Largest infant's share"
  ) |>
  tab_source_note(md(
    "Emotions with no assessable hypothesized configuration are omitted. A single observed configuration can match several emotions, so rows are not additive; see the union total in text. *Largest infant's share* is the proportion of that emotion's total exposure contributed by the single most-exposed infant."
  ))
```

## FIGURE (Phase I)

```{r figure-phase1}
#| fig-width: 6.5
#| fig-height: 5

ff_i  <- phase_totals$valid_fullface[phase_totals$phase == "I"]
ss_i  <- phase_totals$valid_session[phase_totals$phase == "I"]
ratio <- ff_i / ss_i

primary_breaks <- seq(0, 0.15, by = 0.05)

fig_hep <-
  emotion_phase |>
  filter(phase == "I") |>
  mutate(emotion = fct_reorder(str_to_title(emotion), fullface_pct)) |>
  ggplot(aes(x = fullface_pct, y = emotion)) +
  geom_segment(aes(xend = 0, yend = emotion), color = "grey60", linewidth = 0.5) +
  geom_point(size = 2.2) +
  scale_x_continuous(
    name   = "Share of full-face time (person-time)",
    breaks = primary_breaks,
    labels = scales::percent,
    sec.axis = sec_axis(
      transform = ~ .x * ratio,
      name   = "Share of session time",
      breaks = primary_breaks * ratio,
      labels = scales::label_percent(accuracy = 0.01)
    )
  ) +
  labs(y = NULL) +
  theme_bw() +
  theme(
    panel.grid.major.y = element_blank()
  )

fig_hep
ggsave("fig_hep_phase1.png", fig_hep, width = 6.5, height = 5, dpi = 300)
```

## IN-TEXT NUMBERS (Phase I)

Summing across emotions double-counts, because one configuration can match
several. The union below counts each face-second once and is the number to
report as a total.

```{r in-text}
any_hep <-
  occ_events |>
  semi_join(protos_assessable |> select(occ), by = join_by(occ)) |>
  merge_within_face(keys = c("subject", "wave", "phase", "facs_identity")) |>
  summarize(.by = phase, seconds = sum(duration)) |>
  left_join(phase_totals, by = "phase") |>
  mutate(
    pct_fullface = seconds / valid_fullface,
    pct_session  = seconds / valid_session
  )

any_hep |>
  gt() |>
  tab_header(title = "Any hypothesized configuration (union across emotions)") |>
  fmt_number(columns = c(seconds, valid_fullface, valid_session), decimals = 0) |>
  fmt_percent(columns = c(pct_fullface, pct_session), decimals = 2) |>
  cols_label(phase = "Phase", seconds = "Seconds",
             valid_fullface = "Full-face sec", valid_session = "Session sec",
             n_infants = "# Infants",
             pct_fullface = "% Full-face", pct_session = "% Session")

list(
  n_protos              = nrow(proto_status),
  n_complete            = sum(proto_status$status == "complete"),
  n_partial             = sum(proto_status$status == "partial"),
  n_unassessable        = sum(proto_status$status == "unassessable"),
  emotions_unassessable = if (length(emotions_lost)) paste(emotions_lost, collapse = ", ") else "none",
  emotions_assessed     = n_distinct(protos_assessable$emotion),
  hep_pct_fullface_I    = any_hep$pct_fullface[any_hep$phase == "I"],
  hep_pct_session_I     = any_hep$pct_session[any_hep$phase == "I"]
) |>
  enframe(name = "quantity", value = "value") |>
  mutate(value = map_chr(value, \(x) format(unlist(x), digits = 3))) |>
  print(n = Inf)
```

```{r takeaways}
#| output: asis

ff_i   <- any_hep$pct_fullface[any_hep$phase == "I"]
ss_i   <- any_hep$pct_session[any_hep$phase == "I"]
top3   <- emotion_phase |> filter(phase == "I") |> slice_max(fullface_pct, n = 3)
n_none <- emotion_phase |> filter(phase == "I", seconds == 0) |> nrow()

cat(sprintf(
"## Key takeaways (Phase I)

1. **Hypothesized configurations are rare in infants' input.** Infants spent
   %.2f%% of full-face time (%.2f%% of session time) seeing any assessable
   hypothesized configuration.
2. **What exposure there is concentrates in a few emotions.** %s account for most
   of it; %d assessable emotions never occurred at all.
3. **Some emotions could not be assessed.** %d of %d hypothesized configurations
   contain no AU in our coding scheme%s, so their absence would describe the
   scheme rather than infants' input.
4. **Emotion counts are not additive.** Hypothesized sets overlap, so one observed
   configuration can match several emotions. Totals are unions over face-time.
",
  100 * ff_i, 100 * ss_i,
  paste(str_to_title(top3$emotion), collapse = ", "), n_none,
  sum(proto_status$status == "unassessable"), nrow(proto_status),
  if (length(emotions_lost)) paste0(", leaving ", paste(emotions_lost, collapse = " and "),
                                    " with none assessable") else ""
))
```

## SUPPLEMENT

### S1. Phase II

Reported separately, never pooled with Phase I. The caregiver-worn camera in
waves 12 and 15 changes what the infant's view contains, so the two phases are
not directly comparable.

```{r supp-phase2}
n_inf_ii <- phase_totals$n_infants[phase_totals$phase == "II"]

emotion_phase |>
  filter(phase == "II") |>
  arrange(desc(fullface_pct)) |>
  select(emotion, seconds, fullface_pct, session_pct, infants_any, top_infant_share) |>
  mutate(emotion = str_to_title(emotion)) |>
  gt() |>
  tab_header(title = "Phase II: exposure to hypothesized facial configurations") |>
  fmt_number(columns = seconds, decimals = 1) |>
  fmt_percent(columns = c(fullface_pct, session_pct, top_infant_share), decimals = 2) |>
  sub_missing(missing_text = "\u2014") |>
  cols_label(
    emotion = "Emotion", seconds = "Seconds",
    fullface_pct = "% Full-face time", session_pct = "% Session time",
    infants_any = paste0("# Infants (of ", n_inf_ii, ")"),
    top_infant_share = "Largest infant's share"
  )
```

```{r supp-phase-compare}
#| fig-width: 7
#| fig-height: 5

# Both phases side by side, for reference only. Not a test of change over time:
# the recording setup differs between phases.
emotion_phase |>
  mutate(
    phase = factor(phase, levels = c("I", "II"),
                   labels = c("Phase I (6, 9 mo)", "Phase II (12, 15 mo)")),
    emotion = fct_reorder(str_to_title(emotion), fullface_pct, .fun = max)
  ) |>
  ggplot(aes(x = fullface_pct, y = emotion, shape = phase)) +
  geom_line(aes(group = emotion), color = "grey75", linewidth = 2/3) +
  geom_point(size = 2, fill = "white", stroke = 1) +
  scale_shape_manual(values = c(16, 21)) +
  scale_x_continuous(labels = scales::percent) +
  labs(x = "Share of full-face time (person-time)", y = NULL, shape = NULL,
       caption = "Descriptive only: the recording setup differs between phases.") +
  theme_bw() +
  theme(legend.position = "top", panel.grid.minor = element_blank())
```

### S2. Spread across infants

Is an emotion's exposure spread across the sample or driven by a few infants?

```{r supp-spread}
#| fig-width: 7
#| fig-height: 5

emotion_long |>
  filter(phase == "I", duration > 0) |>
  mutate(emotion = fct_reorder(str_to_title(emotion), fullface_pct, .fun = median)) |>
  ggplot(aes(x = fullface_pct, y = emotion)) +
  geom_point(alpha = 0.5, size = 1.6,
             position = position_jitter(height = 0.15, width = 0)) +
  scale_x_continuous(labels = scales::percent) +
  labs(x = "Share of that infant's full-face time", y = NULL,
       caption = "One point per infant-wave with non-zero exposure (Phase I).") +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```

### S3. Files

```{r supp-files}
proto_status |>
  transmute(name, emotion, code = as.character(code), full_aus, kept_aus, reduced, status) |>
  write_csv("supp_hep_S1_assessability.csv")

emotion_long |>
  arrange(phase, emotion, subject, wave) |>
  write_csv("supp_hep_S2_by_infant.csv")

occ_events |>
  inner_join(protos_assessable |> select(emotion, occ, status),
             by = join_by(occ), relationship = "many-to-many") |>
  summarize(
    .by = c(phase, occ, emotion, status),
    n_events  = n_distinct(event_id),
    seconds   = sum(facs_duration),
    n_infants = n_distinct(subject)
  ) |>
  arrange(phase, desc(seconds)) |>
  mutate(occ = as.character(occ)) |>
  write_csv("supp_hep_S3_matching_configs.csv")

write_rds(emotion_long,  "emotion_long.rds")
write_rds(emotion_phase, "emotion_phase.rds")
```
